Designing Types for R, Empirically

### 2020-05-16 paper/main.tex
mono

@@

\begin{figure}[!h]  \centering
  \includegraphics[width=\linewidth]{plots/union_frequency.pdf}
  \caption{Size of unions}\label{fig:unionfreq}
\end{figure}

@@

\begin{figure}[!h]  \centering
  \includegraphics[width=\linewidth]{plots/function_polymorphism.pdf}
  \caption{Function Polymorphism}\label{fig:funpoly}
\end{figure}

Figure~\ref{fig:funpoly} shows the distribution of functions against the number of
polymorphic positions. We observe that \PercMonomorphicFunctions
(\CountMonomorphicFunctions) functions are monomorphic. The remaining
\PercPolymorphicFunctions (\CountPolymorphicFunctions) have at least one
polymorphic parameter or return type. Polymorphic functions are spread across
many packages, \CountMonomorphicPackages out of the 412 packages from our corpus
export only monomorphic functions.

Our type annotation language does not support sophisticated class typing,
and much of the polymorphism in R is due to these classes, namely through
the S3 dispatch mechanism.  In future work, we aim to more fully handle S3
and objects in R.  For now, if we count out

\subsubsection{Discussion}

Now, we aim to enhance the quantitative analysis of our language design by looking at select code snippets from the corpus we analyzed.
We will highlight some of the design decisions that we made, give code examples, and accompany them with function signatures generated by \typetracer.

>>>>

\begin{figure}[!h]  \centering
  \includegraphics[width=.7\linewidth]{plots/union_frequency.pdf}
  \caption{Size of unions}\label{fig:unionfreq}
\end{figure}

@@

@@

\begin{figure}[!h]  \centering
  \includegraphics[width=.7\linewidth]{plots/function_polymorphism.pdf}
  \caption{Function Polymorphism}\label{fig:funpoly}
\end{figure}

If we look at the numbers from the point of view of functions and count how
many of their arguments are polymorphic, we observe that
\PercMonomorphicFunctions (\CountMonomorphicFunctions) functions are
monomorphic. The remaining \PercPolymorphicFunctions
(\CountPolymorphicFunctions) have at least one polymorphic parameter or
return type.  Figure~\ref{fig:funpoly} shows the distribution of functions
against the number of polymorphic arguments. Finally, we count
\CountMonomorphicPackages out of the 412 packages export only monomorphic
functions.

\subsubsection{Discussion}
A number of lessons can be drawn from the data we have gathered.
